Digital Twins for Ocean Robots

### 2024-11-29 paper/paper.tex
final proofreading and edits (part 1)

\keywords{Julia, Ocean, Climate, Robot, Observation, Platform, Sensor, Drifter, Buoy, Profiler, Satellite, Modeling, Artificial Intelligence, Data Assimilation, Parameter Inference, Climatology, Geospatial, Statistics}

@@

DTOR is a framework to access and simulate the global fleet of ocean observing devices (or {\it ocean robots}) that monitor climate change. It brings these observations to Julia and lets us pair ocean robots with virtual counterparts (or {\it twins}). Digital twins provide a bridge to predictive models that enables machine learning. In turn, simulating observations in a digital environment can help evaluate observational strategies a priori, during deployment, or afterwards. In this paper we present the framework, review simulated ocean robots, and discuss envisioned applications.

@@

The DTOR framework was introduced at the Symposium on Advances in Ocean Observation in 2022, and later presented to US-CLIVAR and JuliaCon in 2023. The primary goal of DTOR is to associate ocean observing systems (or {\it ocean robots}) with virtual counterparts (or {\it twins}), and do it for the whole fleet of ocean robots that are currently at sea or have observed the Ocean in the past (Fig.~\ref{fig:ocean-ops}). The virtual twins are to be created through numerical model simulations, which can come in different flavors and languages. The simulation of observing systems in the future, as climate change progresses, is also part of the scope of the DTOR project. At JuliaCon 2023, we presented a solution to (1) access and manipulate the data collected by ocean robots (inc. \pkg{OceanRobots.jl} and \pkg{ArgoData.jl}), and (2) simulate such observations in a digital environment using a hierarchy of models (incl. \pkg{ClimateModels.jl} and \pkg{MITgcm.jl}).

The model hierarchy includes fast climate model emulators, global model output, ocean reanalyses, high-resolution models, and several ways to represent marine ecosystems. Through a streamlined workflow, \pkg{ClimateModels.jl} (JuliaCon 2021, 2023) makes it easy to operate these models that can provide a digital environment for the virtual ocean robot fleet to observe and navigate. \pkg{Drifters.jl} \cite{Forget2021} can be use to predict pathways of ocean robots that tend to follow ocean currents. \pkg{MeshArrays.jl} adds basic geospatial support for global climate model grids, providing capabilities such as interpolation and geolocation on a grid. The framework leverages and links to a variety of highly capable Julia packages from the community. It notably provides extension for visualization with \pkg{Makie.jl}, and tutorial examples in the form of \pkg{Pluto.jl} notebooks.

@@

Part of the fleet of ocean robots currently at sea is depicted in Fig.~\ref{fig:ocean-ops}, focusing on some of the most common observing platforms. To create this map, \pkg{OceanRobots.jl} queries the meta-data-base from Ocean-OPS.org, which keeps track of the whole ocean fleet of scientific observation platforms in real time, and provides a RESTful API. The \pkg{OceanRobots.jl} package provides a simple interface to this API via the `OceanOPS` module. \pkg{OceanRobots.jl} defines data structures to access and utilize data from common ocean robot types such as  `SurfaceDrifter` and `ArgoFloat` (shown in Fig.~\ref{fig:ocean-ops}). A non-exhaustive list of ocean observing platform categories within the scope of \pkg{OceanRobots.jl} is provided in Tab.~\ref{tab:robots}.

@@

A small fraction of the geospatial data obtained in the past using surface ocean drifters (green dots in Fig.~\ref{fig:ocean-ops}) is depicted in Fig.~\ref{fig:drifters} (using \pkg{OceanRobots.jl} and \pkg{Makie.jl}). These drifting buoys tend to follow near-surface ocean currents (at approximately 15 depth). They can measure sea surface temperature and sea level pressure along their trajectory. The data archive currently holds 19396 drifter trajectories, collected over the past 50 years, which allow us to create climatologies such as the one shown in Fig.~\ref{fig:drifters_ve}.

@@

Argo profiling floats (blue dots in Fig.~\ref{fig:ocean-ops}; Code~\ref{lst:ArgoCode}) are one of our main tools to monitor global warming below the sea surface. These devices provide a less detailed view of oceanic pathways (Fig.~\ref{fig:Argo_Float}, top panels) than surface drifters do (e.g., Fig.~\ref{fig:drifters}), since Argo floats only report their location once every ten days. However, Argo floats have a crucial diving capability that surface drifters don't have -- they go up and down the water column to measure temperature and salinity (T,S; bottom panels of Fig.~\ref{fig:Argo_Float}). As a result, the advent of the Argo array in the early 2000s \cite{Roemmich2019} opened up a whole new era of geospatial analysis, state estimation, and parameter inference over the global Ocean  \cite{Forget2007,Forget2010,MFF14,Forget2015a,FFL15,FP15,Forget2024a}. There were 3837 Argo floats at sea on `2024/05/27`, and a total of 18730 in the Argo data base. Since the Argo array is such an important observing system, a dedicated package called \pkg{ArgoData.jl} was created that \pkg{OceanRobots.jl} uses under the hood.

\begin{lstlisting}[
    language = Julia,
%    numbers=left,
    label={lst:ArgoCode},
    caption={Download and visualize one Argo float data as in Fig.~\ref{fig:Argo_Float}.}
]
using OceanRobots, ArgoData, CairoMakie;
argo=read(ArgoFloat(),wmo=6900900);
fig=plot(argo,option=:standard)
\end{lstlisting}

@@

Our ability to simulate the data produced by ocean robots is directly linked to our ability, or lack thereof, to explain observed variations and decipher mechanisms from data. Numerical modeling is thus often motivated and driven by observations. Many research activities involve combining ocean models and observations, with the typical goal of learning model parameters and dynamics from data \cite{Forget2010,Forget2015a,FFL15,FP15,Forget2024a}. Observing system simulations in a digital environment (using models) enable a wide range of applications -- to test out innovative platform or sensor designs ahead of deployment, to optimize global monitoring strategies, or to guide deployed assets in real time for example.

The two simple examples presented below illustrate the simulation of (1) environmental sensors and (2) observing platforms. First, in Fig.~\ref{fig:Argo_Float_simu} we simulate an Argo data collection for temperature (T) and salinity (S) by sampling an ocean climatology along the track of the Argo float (the one from Fig.~\ref{fig:Argo_Float}). This calculation requires (1) a model prediction of $T,S(x,y,z,t)$, which can be based on statistical or mechanistic models on a grid, and (2) tools that deal with the Earth geography, and can localize observations on a grid, and perform interpolation tasks.

To create Fig.~\ref{fig:Argo_Float_simu}, we use \pkg{Climatology.jl} to access the OCCA climatology \cite{Forget2010}, and then interpolate $T,S(x,y,z,t)$ via the geospatial tools in \pkg{MeshArrays.jl} providing. OCCA is essentially a previously trained model that consists of 12 monthly three-dimensional fields of $T,S$ (one per calendar month). Let's note that Figs.~\ref{fig:Argo_Float} and~\ref{fig:Argo_Float_simu} show broadly similar patterns -- this reflects that the predictive model for $T,S$ is skillful and that $T,S$ contrasts see in Fig.~\ref{fig:Argo_Float} largely reflect the $T,S$ sensor moving with the Argo float across the Ocean's geography. Differences between Figs.~\ref{fig:Argo_Float} and~\ref{fig:Argo_Float_simu} can in turn provide a basis for improving the predictive model, for example by including time variability beyond the seasonal cycle, or through various methods for data assimilation, artificial intelligence, or parameter inference \cite{Forget2010,Forget2015a,FFL15,FP15,Forget2024b}.

@@

Our second example focuses on the simulation of observing platforms moving with ocean currents. This basic simulation of surface drifter pathways uses the geostatistical average of velocities calculated in Fig.~\ref{fig:drifters_ve} as a predictive model for the $U,V(x,y,t)$ velocity fields. This gridded data set is simply provided as input to \pkg{Drifters.jl}, which then releases virtual drifters and calculates their trajectories following the flow field (code~\ref{lst:IndDispCode}). The combination of these two model elements represents a previously trained model to predict surface drifter trajectories. While this simple calculation neglects time variability and small scales in ocean currents altogether, the model is sufficient to capture the mean pathway of sea water through this region -- the well-known loop current feeding the Gulf Stream via Florida Strait (Fig.~\ref{fig:drifters_speed_simu}).

@@

using JLD2, CairoMakie, Drifters, OceanRobots;
P=Drifters.Gulf_of_Mexico_setup();
F=FlowFields(u=P.u,v=P.v,period=P.T);
I=Individuals(F,P.x0,P.y0);
[solve!(I,P.T .+P.dT*(n-1)) for n in 1:P.nt]
\end{lstlisting}

>>>>

\keywords{Julia, Ocean, Climate, Robot, Observation, Observing, Platform, Sensor, Drifter, Buoy, Profiler, Model, Modeling, Machine Learning, Artificial Intelligence, Data Assimilation, Parameter Inference, Climatology, Geospatial, Statistics}

@@

DTOR is a framework to access, analyze, and simulate the global fleet of ocean observing devices (or {\it ocean robots}) that monitor climate change. It brings these observations to Julia and lets us pair ocean robots with virtual counterparts (or {\it twins}). Digital twins provide a bridge from the data to predictive models and enable machine learning. In turn, observing system simulations in a digital environment can help evaluate observational strategies a priori, during deployment, or afterwards. In this paper we present the DTOR framework, its supported observing systems, capabilities to simulate ocean robots, and envisioned applications.

@@

The DTOR framework was introduced at the Symposium on Advances in Ocean Observation in 2022, and later presented to US-CLIVAR and JuliaCon in 2023, and the OceanPredict Symposium in 2024. The primary goal of DTOR is to associate ocean observing systems (or {\it ocean robots}) with virtual counterparts (or {\it twins}), and do it for the whole fleet of ocean robots that are currently at sea or have observed the Ocean in the past (Fig.~\ref{fig:ocean-ops}). The virtual twins are to be created through numerical model simulations, which can come in different flavors and languages. The simulation of observing systems in the future, as climate change progresses, is also part of the scope of the DTOR project. At JuliaCon 2023, we presented a solution to (1) access and manipulate the data collected by ocean robots (inc. \pkg{OceanRobots.jl} and \pkg{ArgoData.jl}), and (2) simulate such observations in a digital environment using a hierarchy of models (incl. \pkg{ClimateModels.jl} and \pkg{MITgcm.jl}).

The model hierarchy includes fast climate model emulators, global model output, ocean reanalyses, high-resolution models, and several ways to represent marine ecosystems. Through a streamlined workflow, \pkg{ClimateModels.jl} (JuliaCon 2021, 2023) makes it easy to operate these models that can provide a digital environment for the virtual ocean robot fleet to observe and navigate. \pkg{Drifters.jl} \cite{Forget2021} can be use to predict pathways of ocean robots that tend to follow ocean currents. \pkg{MeshArrays.jl} adds basic geospatial support for global climate model grids, providing capabilities such as interpolation and geolocation on a grid. The framework leverages and links to a variety of highly capable Julia packages from the community. It notably provides user-friendly plotting recipes via its \pkg{Makie.jl} extension, and tutorial examples in the form of \pkg{Pluto.jl} notebooks.

@@

Part of the fleet of ocean robots currently at sea is depicted in Fig.~\ref{fig:ocean-ops}, focusing on a few widely used observing platforms. To create this map, \pkg{OceanRobots.jl} queries the meta-data-base from Ocean-OPS.org, which keeps track of the whole ocean fleet of scientific observing platforms in real time, and provides a RESTful API. The \pkg{OceanRobots.jl} package provides an interface to this API via the `OceanOPS` module. It then defines data structures such as  `SurfaceDrifter` and `ArgoFloat` to access and utilize data from widely used ocean observing platforms. The list of ocean observing platforms currently supported by \pkg{OceanRobots.jl} is in Tab.~\ref{tab:robots}.

@@

A fraction of the geospatial data obtained by surface ocean drifters in the past (green dots in Fig.~\ref{fig:ocean-ops}) is depicted in Fig.~\ref{fig:drifters} using \pkg{OceanRobots.jl} and \pkg{Makie.jl}. These drifting buoys tend to follow ocean currents at approximately 15 depth. They can measure sea surface temperature and sea level pressure along their trajectory. The data archive currently holds 19396 drifter trajectories, collected over the past 50 years, which allow us to create climatologies such as the one shown in Fig.~\ref{fig:drifters_ve}.

@@

Argo profiling floats (blue dots in Fig.~\ref{fig:ocean-ops}; Code~\ref{lst:ArgoCode}) are one of our main tools to monitor global warming below the sea surface. These devices provide a less detailed view of oceanic pathways (Fig.~\ref{fig:Argo_Float}, top panels) than surface drifters do (e.g., Fig.~\ref{fig:drifters}), since Argo floats only report their location once every ten days. However, Argo floats have a crucial diving capability that surface drifters don't have -- they go up and down the water column to measure temperature and salinity (T,S; bottom panels of Fig.~\ref{fig:Argo_Float}). The advent of the Argo array in the early 2000s \cite{Roemmich2019} thus opened up a whole new era of geospatial analysis, state estimation, and parameter inference over the global Ocean  \cite{Forget2007,Forget2010,MFF14,Forget2015a,FFL15,FP15,Forget2024a}. On `2024/05/27` for example there were 3837 Argo floats at sea, and a total of 18730 in the Argo data base. Since the Argo array is such an important observing system, a dedicated package called \pkg{ArgoData.jl} was created that \pkg{OceanRobots.jl} uses under the hood.

\begin{lstlisting}[
    language = Julia,
%    numbers=left,
    label={lst:ArgoCode},
    caption={Download and visualize one Argo float data as in Fig.~\ref{fig:Argo_Float}.}
]
using OceanRobots, CairoMakie;
argo=read(ArgoFloat(),wmo=6900900);
fig=plot(argo,option=:standard)
\end{lstlisting}

@@

Our ability to simulate the data produced by ocean robots is directly linked to our ability, or lack thereof, to explain observed variations and decipher mechanisms from data. Numerical modeling is thus often motivated and driven by observations. Many research activities involve combining ocean models and observations, with the typical goal of learning model parameters, statistics, and dynamics from data \cite{Forget2010,Forget2015a,FFL15,FP15,Forget2024a}. Observing system simulations in a digital environment (using models) enable a wide range of applications -- to test out innovative platform or sensor designs ahead of deployment, to optimize global monitoring strategies, or to guide deployed assets in real time for example.

The two simple examples presented below illustrate the simulation of (1) environmental sensors and (2) observing platforms. First, in Fig.~\ref{fig:Argo_Float_simu} we simulate an Argo data collection for temperature (T) and salinity (S) by sampling an ocean climatology along the track of the Argo float (the one from Fig.~\ref{fig:Argo_Float}). This calculation requires (1) a model prediction of $T(x,y,z,t)$ and $S(x,y,z,t)$, which can be based on statistical or mechanistic models on a grid, and (2) tools that deal with the Earth geography, and can localize observations on a grid, and perform interpolation tasks.

To create Fig.~\ref{fig:Argo_Float_simu}, we use \pkg{Climatology.jl} to access the OCCA climatology \cite{Forget2010}, and then interpolate $T(x,y,z,t)$, $S(x,y,z,t)$ via geospatial tools from \pkg{MeshArrays.jl}. The OCCA climatology is a previously trained model that consists of 12 monthly three-dimensional fields of $T,S$ (one per calendar month). The fact that Figs.~\ref{fig:Argo_Float} and~\ref{fig:Argo_Float_simu} show broadly similar patterns reflects that the predictive model is skillful, and that $T,S$ contrasts seen in Fig.~\ref{fig:Argo_Float} largely result from the sensor moving with the Argo float across the Ocean's $T,S$ geography. Differences between Figs.~\ref{fig:Argo_Float} and~\ref{fig:Argo_Float_simu} can in turn provide a basis for improving the predictive model, for example by including time variability beyond the seasonal cycle, or through various methods for data assimilation, artificial intelligence, or parameter inference \cite{Forget2010,Forget2015a,FFL15,FP15,Forget2024b}.

@@

Our second example focuses on the simulation of observing platforms moving with ocean currents. This basic simulation of surface drifter pathways uses the geostatistical average of velocities calculated in Fig.~\ref{fig:drifters_ve} as a predictive model for the $U(x,y,t)$, $V(x,y,t)$ velocity fields. In Code~\ref{lst:IndDispCode}, this gridded data set is provided as input to \pkg{Drifters.jl}, which then releases virtual drifters and calculates their trajectories following the flow field (Fig.~\ref{fig:drifters_speed_simu}). While this simple drifter trajectory prediction model neglects stochastic variability and small scales in ocean currents altogether (Fig.~\ref{fig:km_scale}), it is sufficient to capture the main mean pathway of sea water through the Gulf of Mexico region -- the well-known loop current feeding the Gulf Stream via Florida Strait (Fig.~\ref{fig:drifters_speed_simu}).

@@

using OceanRobots, Drifters, CairoMakie ;
P=Drifters.Gulf_of_Mexico_setup();
F=FlowFields(u=P.u,v=P.v,period=P.T);
I=Individuals(F,P.x0,P.y0);
[solve!(I,P.T .+P.dT*(n-1)) for n in 1:P.nt];
\end{lstlisting}
